Experience Report: Distributed Computing for the Common Man

### 2010-03-28 doc/icfp-2010/main.tex
some corrections as suggested by Claude [KK]

\authorinfo{Jean-Christophe Filli\^{a}tre}
           {CNRS, LRI, Univ Paris-Sud, Orsay F-91405\\
            INRIA Saclay - Île-de-France, ProVal, Orsay, F-91893}
           {filliatr@lri.fr}
\authorinfo{Kalyan Krishnamani}
           {INRIA Saclay - Île-de-France, ProVal, Orsay, F-91893}
           {kalyan@lri.fr}

@@

\begin{abstract}
  In this paper, we present a distributed computing library for
  functional programming languages. The main features of this library
  include (1) a highly-generic API, (2) several implementations to
  adapt to different computing environments such as sequential,
  multi-core or network, and (3) a reliable fault tolerance mechanism.
  This paper describes the motivation behind this work, as well as
  the design and implementation of the library. It also demonstrates
  the potential of the library with realistic experiments.
\end{abstract}

@@

This paper introduces a generic distributed computing library. This
work was initially motivated by the computing needs that exist in our
own research team. Our applications include large-scale deductive
program verification, which amounts to checking the validity of a
large number of logical formulas using a variety of automated theorem
provers~\cite{filliatre07cav}. Our computing infrastructure consists
of a few powerful multi-core machines (typically 8 to 16 cores) and
several desktop PCs (typically dual-core). However, for our
application needs, there is no library that helps in exploiting such a
computing infrastructure in our favorite functional programming
language.  Hence we designed and implemented such a
library\footnote{This library is implemented in \Ocaml, but the
  implementation should be straightforward in any functional
  programming language.}, which is the subject of this paper.

The distributing computing library presented in this paper is not a
library that helps in parallelizing computations. Rather, it provides
facilities for reliable, distributed execution of parallelizable
computations. In particular, it provides a set of APIs that allows the
execution of large-scale parallelizable computations, very relevant to
our application needs (and presumably relevant to a variety of
real-world applications), over multiple cores in the same machine or
over a network of machines. The most important features of our library
are the following:
\begin{itemize}
\item \emph{genericity}: 
  it allows various patterns of polymorphic computations;
\item \emph{simplicity}: switching between multiple cores on the same
  machine and a network of machines is as simple as changing a couple
  of lines of code;
\item \emph{task distribution and fault-tolerance}: 
  it relieves the user from implementing task distribution routines
  and reliable fault-tolerance mechanisms.
\end{itemize}
It is worth noting that the library is not targeted at applications
running on server farms, crunching enormous amounts of data. Neither
is it limited to research labs possessing a few number of powerful
computing infrastructure. It serves a variety of users and a wide
spectrum of needs, from desktop PCs to networks of machines, and hence
the title.

>>>>

\authorinfo{Jean-Christophe Filli\^{a}tre}
           {CNRS, LRI, Univ Paris-Sud, Orsay F-91405\\
            INRIA Saclay - \^{I}le-de-France, ProVal, Orsay, F-91893}
           {filliatr@lri.fr}
\authorinfo{Kalyan Krishnamani}
           {INRIA Saclay - \^{I}le-de-France, ProVal, Orsay, F-91893}
           {kalyan@inria.fr}

@@

\begin{abstract}
  We present a distributed computing library for
  functional programming languages. The main features of this library
  include (1) a highly-generic API, (2) several implementations to
  adapt to different computing environments such as sequential,
  multi-core or network, and (3) a reliable fault tolerance mechanism.
  This paper describes the motivation behind this work, as well as
  the design and implementation of the library. It also demonstrates
  the potential of the library with realistic experiments.
\end{abstract}

@@

This paper introduces a generic library for distributed computing. This
work was initially motivated by the computing needs that exist in our
own research team. Our applications include large-scale deductive
program verification, which amounts to checking the validity of a
large number of logical formulas using a variety of automated theorem
provers~\cite{filliatre07cav}. Our computing infrastructure consists
of a few powerful multi-core machines (typically 8 to 16 cores) and
several desktop PCs (typically dual-core). However, for our
application needs, there is no library that helps in exploiting such a
computing infrastructure in our favorite functional programming
language.  Hence we designed and implemented such a
library\footnote{This library is implemented in \Ocaml, but the
  implementation should be straightforward in any functional
  programming language.}, which is the subject of this paper.

The distributed computing library presented in this paper is not a
library that helps in parallelizing computations. Rather, it provides
facilities for reliable, distributed execution of parallelizable
computations. In particular, it provides a set of APIs that allows the
execution of large-scale parallelizable computations, very relevant to
our application needs (and presumably relevant to a variety of
real-world applications), over multiple cores in the same machine or
over a network of machines. The most important features of our library
are the following:
\begin{itemize}
\item \emph{genericity}: 
  it allows various patterns of polymorphic computations;
\item \emph{simplicity}: switching between multiple cores on the same
  machine and a network of machines is as simple as changing a couple
  of lines of code;
\item \emph{task distribution and fault-tolerance}: 
  it relieves the user from implementing task distribution routines
  and reliable fault-tolerance mechanisms.
\end{itemize}
It is worth noting that the library is not targeted at applications
running on server farms, crunching enormous amounts of data. Neither
is it limited to research labs possessing a few number of powerful
computing infrastructure. It serves a variety of users and a wide
spectrum of needs, from desktop PCs to networks of machines, and hence
the title.
